---
title: "Getting started with GLBFP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with GLBFP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(2026)
```

This vignette shows the basic workflow for pointwise and grid-based density
estimation with `GLBFP`.

## Installation

```{r, eval = FALSE}
install.packages("remotes")
remotes::install_github("AurelienNicosiaULaval/GLBFP")
```

## Load the package

```{r}
library(GLBFP)
```

## Estimate density at one point

The functions `ASH()`, `LBFP()`, and `GLBFP()` estimate the density at a single
point. The data are supplied as a numeric matrix or data frame with observations
in rows.

```{r}
x <- matrix(rnorm(300), ncol = 1)
b <- compute_bi_optim(x, m = 1)

fit <- GLBFP(x = 0, data = x, b = b, m = 1)
fit
```

The returned object contains the evaluation point, the estimated density, the
bandwidth, and the shift parameter. For `LBFP()` and `GLBFP()`, uncertainty
components are also returned.

```{r}
names(fit)
summary(fit)
predict(fit)
```

## Estimate density on a grid

The `*_estimate()` functions evaluate the same estimator over a regular grid or
a user-supplied set of points.

```{r}
grid_fit <- GLBFP_estimate(data = x, b = b, m = 1, grid_size = 80)
head(cbind(grid_fit$grid, density = grid_fit$densities))
```

For one-dimensional grids, the plot method returns a `ggplot2` object.

```{r}
plot(grid_fit)
```

## Using the included data

The `ashua` data contain daily flow and level observations for the Ashuapmushuan
river. The example below uses a small grid so the vignette remains fast.

```{r}
data("ashua")

river_data <- ashua[, c("flow", "level")]
b2 <- c(8, 0.4)
x0 <- c(mean(river_data$flow), mean(river_data$level))

fit2 <- GLBFP(x = x0, data = river_data, b = b2, m = c(1, 1))
fit2

grid_fit2 <- GLBFP_estimate(
  data = river_data,
  b = b2,
  m = c(1, 1),
  grid_size = 15
)

plot(grid_fit2, contour = TRUE)
```

## Input expectations

The package expects finite numeric data. Missing values should be removed or
imputed before estimation. Constant data require explicit non-degenerate
`min_vals` and `max_vals` bounds because a density estimator for continuous data
needs a positive evaluation range.
