---
title: "Choosing between ASH, LBFP and GLBFP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Choosing between ASH, LBFP and GLBFP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(20260607)
```

The package exposes three related grid-count density estimators.

- `ASH()` and `ASH_estimate()` implement averaged shifted histogram estimates.
- `LBFP()` and `LBFP_estimate()` implement linear blend frequency polygon estimates.
- `GLBFP()` and `GLBFP_estimate()` implement the general linear blend frequency polygon estimate.

All estimators share the same basic arguments: `data`, `b`, optional grid bounds,
and for ASH/GLBFP the shift vector `m`.

```{r}
library(GLBFP)

x <- cbind(rnorm(200), rnorm(200, sd = 1.25))
b <- c(0.75, 0.9)
m <- c(2, 2)
point <- c(0, 0)

fits <- list(
  ASH = ash(point, x, b = b, m = m),
  LBFP = lbfp(point, x, b = b),
  GLBFP = glbfp(point, x, b = b, m = m)
)

vapply(fits, function(z) z$estimation, numeric(1))
```

Grid estimates can be compared through a common data-frame interface.

```{r}
grid_ash <- ash_estimate(x, b = b, m = m, grid_size = 15)
grid_lbfp <- lbfp_estimate(x, b = b, grid_size = 15)
grid_glbfp <- glbfp_estimate(x, b = b, m = m, grid_size = 15)

comparison <- data.frame(
  method = c("ASH", "LBFP", "GLBFP"),
  mean_density = c(
    mean(grid_ash$densities),
    mean(grid_lbfp$densities),
    mean(grid_glbfp$densities)
  ),
  max_density = c(
    max(grid_ash$densities),
    max(grid_lbfp$densities),
    max(grid_glbfp$densities)
  )
)

comparison
```

As a practical starting rule:

- use `LBFP` when `m = 1` is sufficient and a simple linear estimator is desired;
- use `GLBFP` when a tunable shifted estimator is desired;
- use `ASH` when an averaged shifted histogram representation is preferred.

The bandwidth vector `b` usually matters more than small changes in `m`.
Use `compute_bi_optim()` as a reproducible starting point, then inspect
sensitivity around that value.
