---
title: "Fast sparse-prefix computation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fast sparse-prefix computation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
set.seed(20260607)
```

`GLBFP` uses a sparse-prefix traversal for finite-stencil grid-count density
estimators. The estimator is unchanged. The computation is accelerated by
counting occupied grid cells once and pruning stencil branches that cannot
reach an occupied cell.

```{r}
library(GLBFP)

x <- cbind(rnorm(250), rnorm(250), rnorm(250))
b <- c(0.8, 0.8, 0.8)
m <- c(2, 2, 2)

fit <- glbfp_estimate(x, b = b, m = m, grid_size = 7)

c(
  grid_points = nrow(fit$grid),
  nominal_stencil = prod(2 * m),
  median_visited = median(fit$visited),
  median_prefix_nodes = median(fit$prefix_nodes)
)
```

The `visited` field records the number of nonzero occupied cells reached for
each evaluation point. The `prefix_nodes` field records the number of explored
prefix nodes. These diagnostics are useful for checking whether sparsity is
helping in a given data set.

```{r}
summary(fit)
```

The sparse traversal also powers `ASH_estimate()` and `LBFP_estimate()`.

```{r}
ash_fit <- ash_estimate(x, b = b, m = m, grid_size = 7)
lbfp_fit <- lbfp_estimate(x, b = b, grid_size = 7)

rbind(
  ASH = c(median_visited = median(ash_fit$visited), max_visited = max(ash_fit$visited)),
  LBFP = c(median_visited = median(lbfp_fit$visited), max_visited = max(lbfp_fit$visited)),
  GLBFP = c(median_visited = median(fit$visited), max_visited = max(fit$visited))
)
```

The sparse-prefix implementation is written in R for CRAN portability. It is
adapted from the finite-stencil sparse traversal project used for the package
development experiments.
